---
title: "Factor Analysis Model"
---

```{r}
#| message: false
#| warning: false

# Packages Needed
library(lavaan)
```

Factor analysis models an unobserved continuous latent variable $\xi$, based on continuous observed items $X_1, X_2, \dots$. Each item is connected with the factor $\xi$ by a **linear model**:

$$
\begin{align}
X_1 & = \tau_1 + \blue{\lambda_1} \ \xi + \delta_1 \\
X_2 & = \tau_2 + \blue{\lambda_2} \ \xi + \delta_2 \\
X_3 & = \tau_3 + \blue{\lambda_3} \ \xi + \delta_3 \\
\end{align}
$$

::: small
$\tau$ is the intercept of the model. $\lambda$ is the coefficient, called the factor loadings. $\delta$ is the error term.
:::

```{r}
#| message: false
#| warning: false
#| echo: false

df <- readRDS("ESS4values.rds")
```

```{r}
#| message: false
#| warning: false
#| comment: "#>"
#| class-output: r
#| results: hold

formula <- '
factor =~ rich + admire + success + respect
'
# Formula is always factor =~ item1 + item 2
# You can add multiple factors f1, f2, and multiple measurement models together

model <- sem(
  formula,
  data = df,
  std.ov= TRUE, #standardises the observed variables
  std.lv= TRUE,
  missing = "fiml")
      # this allows use of missing data, but may take longer
coef(model)
    # You can also use summary(model), which gives a more comprehensive output
```

::: small
The factor=\~item are $\lambda$. The item\~\~item are $\delta$. The item\~1 are $\tau$. Each column is one specific model for one item.
:::

The factor loadings $\blue{\lambda_i}$ represent the relationship/**covariance** between any item and a factor. These factor loadings help us interpret our latent variable. The sign of the factor loading tells us the direction in which our latent variable is measuring. The absolute size of the factor loading tells us how important that item is to the factor.

::: small
In the output above, all $\lambda$ are positive, meaning as item increases, the factor also increases. *success* is the item with the largest $\lambda$, so it is measured the most in the factor.
:::

**Factor scores** $\widehat\xi$ are essentially values of the latent variable for individuals in our study. This allows us to use our observed items to calculate the latent variable value that any individual should have, which we can then put into another statistical model. We can calculate them in R:

```{r}
#| warning: false
#| message: false
#| comment: "#>"
#| class-output: r

factor.scores <- lavPredict(model)

# visualise factor scores (not required)
head(factor.scores)
```

::: small
These are the first 6 observation's factor scores. You can use the variable you saved the factor scores to in a regression or other statistical model.
:::
