---
title: "Lagged Response Models"
---

A **lagged response model** is a regression model that says that the outcome variable not only depends on the input variables, but also previous values of the outcome from the past.

$$
\eta_{it} = \beta_0 + \underbrace{\beta_1 X_{1,it} + \beta_2 X_{2,it}}_{t} + \underbrace{\beta_3 \ \eta_{i,t-1}}_{t-1} + \eps_{it}
$$

::: center-graph
```{dot}
//| fig-width: 6
//| fig-height: 1.2
digraph example2 {
  bgcolor="transparent";
  // Nodes
  Y1 [shape=box, label=<<FONT FACE="Helvetica">&eta; (t-1)</FONT>>]
  X1 [shape=box, pos = "2,0!", label="X1 and X2"]
  Y2 [shape=box, label=<<FONT FACE="Helvetica">&eta; (t)</FONT>>]

  // Edges
  {X1 -> Y1}
  {X1 -> Y2}
  {rank=same; Y1 -> Y2 [label="Lag"]}
  
  graph [nodesep=0.5, ranksep=0.5]

}
```
:::

Recall the response variable is itself a latent variable, measured with observed items. This also implies a measurement model relating the latent response with each observed item:

$$
\b X_{t} = \b\tau_t + \b\eta'\b\lambda_t + \b\delta_{t}
$$

::: small
This is just the model for one (of many) items. Note how the parameters have the subscript $t$ - this implies a different measurement model for each item, at each time period.
:::

However, because each item is observed over multiple time periods, it is likely that an item from $t=1$ is correlated with the same item from $t=2$ ( autocorrelation). We have to take this into account in one of two ways:

1.  We could **constrain** the model, assuming that the measurement model is identical throughout different time periods. That means $\tau_t$ and $\lambda_t$ will now no longer depend on the time period (the same for all time periods, for the same item).
2.  Or we could complicate the model, by including correlation between the error terms $\eps$ for the same item at different times.

<br />

To implement lagged response models, we need the **lavaan** package:

```{r, eval = F}
library(lavaan)
```

For a constrained two-period model, the estimation is as follows:

```{r, eval = F}
model <- '
# explanatory variables Measurement models 
  explanatory1 =~ NA*X1 + X2 + X3
  explanatory2 =~ NA*Z1 + Z2 + Z3

# response variables for time = 1 and time = 2
	response1 =~ la*Y1t1 + lb*Y2t1 + lc*Y3t1
	response2 =~ la*Y1t2 + lb*Y2t2 + lc*Y3t2

# time = 1 model
  response1 ~ explanatory1 + explanatory2
  
# time = 2 model with lagged time = 1
  response2 ~ explanatory1 + explanatory2 + response1

# Fixing the variances of the exogenous latent variables LR and AL  
  LR~~1*LR
  AL~~1*AL
'

sem <- sem(model, data = my_data, missing="FIML")
summary(sem)
```

::: small
The labels la\*, lb\*, and lc\* indicate to R to make all factor loadings with the same labels equivalent to each other. Notice how we put the same labels on the different time period measurement models.
:::

For a unconstrained model, the code is as follows:

```{r, eval = F}
model <- '
# explanatory variables Measurement models 
  explanatory1 =~ NA*X1 + X2 + X3
  explanatory2 =~ NA*Z1 + Z2 + Z3

# response variables for time = 1 and time = 2
	response1 =~ Y1t1 + Y2t1 + Y3t1
	response2 =~ Y1t2 + Y2t2 + Y3t2

## Error covariance in the measurement model
  Y3t1 ~~ Y3t2

# time = 1 model
  response1 ~ explanatory1 + explanatory2
  
# time = 2 model with lagged time = 1
  response2 ~ explanatory1 + explanatory2 + response1

# Fixing the variances of the exogenous latent variables LR and AL  
  LR~~1*LR
  AL~~1*AL
'

sem <- sem(model, data = my_data, missing="FIML")
summary(sem)
```
