---
title: "Reliability and Factor Scores"
---

::: small
Make sure you have read the previous page on [factor analysis](factor.qmd) before this.
:::

Recall our measurement models linking each item $i$ to our factor:

$$
X_i = \tau_i + \lambda_i\ \xi + \delta_i
$$

The the error term has a mean of 0 and a variance of $\theta_i$. The variance of each item can be shown to equal:

$$
Var(X_i) = \lambda_i^2 + \theta_i
$$

Thus, this allows us to essentially "split" the variance in any item into two parts:

1.  $\lambda_i^2$ is the part of the variance in the item explained by the latent factor. We call this the **communality** of the item $i$.
2.  $\theta_i$ is the **residual variance**, the part of the variance **not** explained by our factor.

We can also calculate the percentage/proportion of variance in item $i$ that our factor explains, called the **reliability**:

$$
\text{Reliability} = \frac{\lambda_i^2}{Var(X_i)} = \frac{\lambda_i^2}{\lambda_i^2+\theta_i}
$$

::: small
If all items are standarised to a standard normal, then $\lambda_i^2$ is equal to the reliability.
:::

Items with higher reliability are considered more "accurate" measures of the latent variable factor. They thus ensure better model estimation - and when a factor has low relaibility, we will often drop it.

**Factor scores** $\widehat\xi$ are essentially values of the latent variable for individuals in our study. This allows us to use our observed items to calculate the latent variable value that any individual should have, which we can then put into another statistical model.

$$
\widehat\xi = \omega_0 + \omega_1 X_1 + \omega_2 X_2 + \dots
$$

The weights are calculated based on the communalities. The items with the highest communalities tend to get the strongest weights, while the items with the least communalities get the smallest weights.

<br />

To calculate factor analysis, we have to first run a factor analysis model in the same way we did in the [last page](factor.qmd).

```{r, eval = FALSE}
library(psych)
library(GPArotation)

# eliminate missing observations
all.obs <- apply(my_data, 1, FUN=function(x){all(!is.na(x))})
dta <- my_data[all.obs,]

# factor model
fa <- fa(data[,items], nfactors=1, fm="ml")
```

R will automatically calculate factor scores in the estimation process, so all we have to do is access it within our output object:

```{r, eval = F}
fa$scores
```

You can save this into your dataset, and use for other purposes.
