---
title: "Structural Equation Models"
---

```{r}
#| message: false
#| warning: false

# Packages Needed
library(lavaan)
```

Structural Equation Models (SEMs) allow us to combine latent models with linear regression models between the latent variables.

::: center-graph
```{dot}
//| fig-width: 6
//| fig-height: 1.2
digraph example2 {
  bgcolor="transparent";
  // Nodes
  F1 [shape=box, label=<<FONT FACE="Helvetica">&xi; (Unobserved)</FONT>>]
  X1 [shape=box, pos = "2,0!", label="X1"]
  X2 [shape=box, pos = "1,-2!", label="X2"]
  X3 [shape=box, pos="1,0!", label="X3"]
  F2 [shape=box, label=<<FONT FACE="Helvetica">&eta; (Unobserved)</FONT>>]
  X4 [shape=box, pos = "2,0!", label="Y1"]
  X5 [shape=box, pos = "1,-2!", label="Y2"]
  X6 [shape=box, pos="1,0!", label="Y3"]

  // Edges
  {F1 -> X1 [label=<<FONT FACE="Arial">&lambda;</FONT>>]}
  {F1 -> X2 [label=<<FONT FACE="Arial">&lambda;</FONT>>]}
  {F1 -> X3 [label=<<FONT FACE="Arial">&lambda;</FONT>>]}
  {F2 -> X4 [label=<<FONT FACE="Arial">&lambda;</FONT>>]}
  {F2 -> X5 [label=<<FONT FACE="Arial">&lambda;</FONT>>]}
  {F2 -> X6 [label=<<FONT FACE="Arial">&lambda;</FONT>>]}
  {rank=same; F1 -> F2 [label="Linear Model"]}
  
  graph [nodesep=0.5, ranksep=0.5]

}
```
:::

This structural equation model contains a linear regression between the latent explanatory variable explaining the latent outcome variable.

$$
\eta = \beta_0 + \beta_1\ \xi + \eps
$$

Both $\eta$ and $\xi$ are latent variables, which need to be estimated with a latent measurement model like factor analysis. We can estimate all measurement models and structural regressions together in a structural equation model:

```{r}
#| message: false
#| warning: false
#| echo: false

df <- readRDS("ESS4values.rds")
```

```{r}
#| message: false
#| warning: false
#| comment: "#>"
#| class-output: r
#| results: hold

formula <- '
# 2 Measurement Models
  factor1 =~ rich + admire + success + respect
  factor2 =~ friend + equal + nature + care

# Stuctural regression
  factor1 ~ factor2
'

model <- sem(
  formula,
  data = df,
  std.ov= TRUE, #standardises the observed variables
  std.lv= TRUE,
  missing = "fiml")
      # this allows use of missing data, but may take longer
coef(model)
    # You can also use summary(model), which gives a more comprehensive output
```

::: small
See [factor analysis](factor.qmd) on how most terms are interpreted. factor1\~factor2 is the coefficient $\beta_1$ of the structural regression.
:::

The interpretations of the model are quite straight forward - the measurment models are interpreted in the same way as factor analysis, and the linear model between latent variable is interpreted the same way as a linear model between any other variables.

::: small
Ex: the linear model says that as explanatory factor increases by 1, response increases by an expected $\beta$ units.
:::

Structural can be far more complicated. For example:

::: center-graph
```{dot}
//| fig-width: 5
//| fig-height: 1.4
digraph example2 {
  bgcolor="transparent";
  // Nodes
  X1 [shape=box, label=<<FONT FACE="Helvetica">&xi; 1</FONT>>]
  X2 [shape=box, label=<<FONT FACE="Helvetica">&xi; 2</FONT>>]
  Y1 [shape=box, label=<<FONT FACE="Helvetica">&eta; 2</FONT>>]
  Y2 [shape=box, label=<<FONT FACE="Helvetica">&eta; 2</FONT>>]

  // Edges
  {rank=same; X1 -> Y1 [label=<<FONT FACE="Arial">&beta;1</FONT>>]}
  {rank=same; X2 -> Y2 [label=<<FONT FACE="Arial">&gamma;2</FONT>>]}
  X1 -> Y2 [label=<<FONT FACE="Arial">&gamma;1</FONT>>]
  X2 -> Y1 [label=<<FONT FACE="Arial">&beta;2</FONT>>]
  X1 -> X2 [dir=both, label = "Correlation"]
  Y1 -> Y2 [dir=both, label = "Correlation"]
  
  graph [nodesep=0.5, ranksep=0.5]

}
```
:::

::: small
Note: we cannot have linear models between two independent, or two dependent variables. They have to be seperate - within each group, only correlations are possible.
:::

In this example, we are measuring the correlation between the independent variables, the correlation between the dependent variables, and we have two regression models:

$$
\begin{align}
\eta_1 & = \beta_0 + \beta_1\ \xi_1 + \beta_2\ \xi_2 + \eps_1 \\
\eta_2 & = \gamma_0 + \gamma_1\ \xi_1 + \gamma_2\ \xi_2 + \eps_2 \\
\end{align}
$$

To implement these more complicated SEMs in R, we just adjust the formula input. Below shows different examples of how we can specify structural models.

```{r}
#| eval: false

formula <- '
# Measurement Models, same format, see above.

# Stuctural regression with multiple explanatory variables
  factor1 ~ factor2 + factor3

# structural covariances
  factor1 ~~ factor2
'

```

::: small
The estimation is the same - use the sem() command.
:::
