---
title: "Callaway and Sant'Anna Flexible Matching and Reweighting Estimator"
---

Callaway and Sant'Anna (2021) proposes to solve the [issues with two-way fixed effects](twfestaggered.qmd) by carefully selecting the comparisons, and properly weighting the comparisons:

1.  We first divide all units $i$ into groups, based on their first initial year of receiving treatment. Each unit who was first treated in year $g$ is assigned to group $g$.

2.  Then, we define a new causal estimand of interest: what they call the group-time ATT. This is essentially the dynamic treatment effect of group $g$ at time $t$:

$$
\tau_{g, t}^\text{ATT} = \green{\mean Y_t^{(1)}} - \red{\mean Y_t^{(0)}} \ | \ G_{i, g} = 1
$$

::: small
Where $G_{i,g}$ is a variable that equals 1 if unit $i$ is initially treated in group $g$. It is 0 otherwise.
:::

3.  Now the issue (as is the case for all DiD), is that the counterfactual $\red{\mean Y_t^{(0)}}$ is unobserved. Now, we need some way to estimate it.

4.  To estimate the "trend" if the treated units were hypothetically untreated, we look only at the **never-treated** group. This ensures that we are making proper comparisons (instead of comparing with already-treated units). We use this group to approximate the trend if hypothetically our treated units were untreated.

::: small
If this estimation of "trend" makes little sense, see the [intro page on DiD](did.qmd).
:::

5.  With the trend, we can now estimate the missing counterfactual, allowing us to obtain the group-time ATT $\tau_{g,t}^\text{ATT}$.

6.  Now, we weight each $\tau_{g,t}^\text{ATT}$ together based on how frequent each group $g$ is. By properly weighting, we solve the weighting issue of TWFE.

::: small
The final step of weighting can be done to obtain one overall ATT, or dynamic ATTs, or even we can decide not to aggregate and report each individual group-time $\tau_{g,t}^\text{ATT}$.
:::

Unlike the other estimators covered, Callaway and Sant'Anna do not rely on regression models. The comparisons are made through grouping and matching, and the individual effects are weighted.

<br />

To implement this estimator, we will need the **did** package:

::: small
Note: there are issues when including covariates, so this method may not be the best if you need covariates to condition for parallel trends.
:::

```{r, eval = F}
library(did)
```

The data will need pre-processing. We will need to make sure the never-treated group has a initial treatment year of 0, and we will need the unit variable to be a numerical id variable.

```{r, eval = F}
dta <- my_data 

# set treat_year = 0 for never-treated (assuming NA)
dta$treat_year <- ifelse(is.na(dta_modify$treat_year), 0, dta_modify$treat_year)

# numerical id variable
dta <- dta %>%
  group_by(unit) %>%
  mutate(id = cur_group_id()) %>%
  ungroup()
```

Now, we can use the pre-process function in the **did** package to make the final touches before starting.

```{r, eval = F}
set.seed(17)
dta <- dta %>%
  pre_process_did(yname = "outcome",
                  tname = "treatment",
                  idname = "id", #don't change
                  gname = "treat_year",
                  allow_unbalanced_panel = T,
                  data = .)
```

Now, we can run the analysis:

```{r, eval = F}
model <- dta[['data']] %>%
  att_gt(yname = "outcome",
         tname = "year",
         idname = "id", #don't change
         gname = "treat_year",
         control_group = c("nevertreated"),
         est_method = 'dr',
         base_period = 'universal',
         allow_unbalanced_panel = T,
         data = .)




```

We can view our results in multiple ways. For a singular ATT:

```{r, eval = F}
att <- did::aggte(model, type = "simple", na.rm = TRUE)
summary(att)
```

We can also see ATT's grouped by initial treatment year:

```{r, eval = F}
att_group <- aggte(model, type = "group", na.rm = TRUE)
ggdid(att_group)
```
