---
title: "Types of DiD Designs"
---

```{r}
#| warning: false
#| message: false

# Packages Needed
library(panelView)
```

Difference-in-differences designs can vary based on two factors: the **timing** of treatment implementation, and the **absorption** of treatment.

The timing of treatment implementation is about when units receive the treatment. There are two types of treatment timing:

|                        |                                                                                                                     |
|----------------------|-------------------------------------------------|
| [Non-Staggered]{.mark} | All units who ever receive the treatment, all begin the treatment in the same year.                                 |
| [Staggered]{.mark}     | Units begin the treatment in different years. For example, some units begin treatment in 2003, others in 2004, etc. |

: {tbl-colwidths="\[35,65\]" .bordered}

::: small
The non-staggered design is often called the classical design.
:::

The absorption of treatment is whether or not any unit who receives the treatment later on reverses/loses the treatment.

|                        |                                                                                      |
|-----------------------|--------------------------------------------------|
| [Absorbing]{.mark}     | All units who ever receive the treatment, remain treated until the end of the study. |
| [Non-Absorbing]{.mark} | Some units who receive the treatment will leave/reverse the treatment later.         |

: {tbl-colwidths="\[35,65\]" .bordered}

::: small
The research into non-absorbing DiD is quite limited as of now.
:::

The type of study is important for the estimator we will need. We can determine the type of our study by using the **panelView** package.

```{r}
#| echo: false

data(base_did, package = "fixest")
df <- base_did
colnames(df) <- c("outcome", "covariate", "unit", "time", "post", "ever.treat")
df$treatment <- df$post*df$ever.treat
```

```{r}
#| fig-height: 3

panelview(
  # required options
  data    = df,                    # replace with your data
  formula = outcome ~ treatment,   # replace with outcome and treat var
  index   = c("unit", "time"),     # replace with unit var and time var
  gridOff = T,
  
  # aesthetic options
  xlab       = "Time Periods",        # custom x axis label
  ylab       = "Unit",                # custom y axis label
  main       = "Treatment Adoption",  # custom title name
  cex.main   = 11,                    # font size of title
  cex.lab    = 10,                    # font size of labels
  cex.axis.x = 8,                     # font size of x axis
  cex.axis.y = 3,                     # font size of y axis
  background = "white"                # background color
)

```

Above we can see this study is a non-staggered and absorbing design. We can tell this because all the blue (under treatment) begins at the same time period, and no unit that gets treated becomes untreated in a later time period.

It is quite conventional to include a diagram like the above PanelView during DiD studies to help visualise the treatment adoption over time.
