---
title: "Difference-in-Differences"
---

::: small
This example is from [Card and Kreuger (1993)](https://www.nber.org/papers/w4509) with modified numbers.
:::

In April 1992, New Jersey increased its minimum wage. We might wonder how that caused changes in employment. We have data from before (Mar 1992) and after (Dec 1992):

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(fixest)
df.nj <- data.frame(
  time = c(0, 1),
  nj = c(20, 24)
)

df.pa <- data.frame(
  time = c(0, 1),
  pa = c(18, 19)
)

df.pot <- data.frame(
  time = c(0, 1),
  nj.pot = c(20, 21)
)

```

```{r}
#| fig-height: 2
#| echo: false

ggplot() +
  geom_point(aes(x = time, y = nj, color = "col2"), data = df.nj) +
  geom_line(aes(x = time, y = nj, color = "col2"), data = df.nj) +
  scale_color_manual(
    name = "Legend",
    values = c("col2" = "darkgreen"),
    labels = c("New Jersey (Treated - Min Wage Increase)")
  ) +
  theme_bw() +
  scale_x_continuous(breaks = c(0, 1), labels = c("Mar 1992","Dec 1992")) +
  xlab(NULL) + ylab("Employment Rate") +
  ylim(17, 27)

```

We **cannot** calculate the treatment effect. This is because the actual treatment effect, according to potential outcomes, requires the counterfactual outcome if New Jersey had hypothetically not received the treatment in Dec 1992.

::: small
The actual treatment effect would be employment rate in Dec 1992 with min wage increase, minus the employment rate in Dec 1992 without the min wage increase.
:::

Of course, we do not observe the hypothetical counterfactual of New Jersey not undergoing the minimum wage increase in Dec 1992. However, what we can do is look at another state that **did not undergo the treatment** (untreated), like Pennsylvania.

```{r}
#| fig-height: 2
#| echo: false

ggplot() +
  geom_point(aes(x = time, y = nj, color = "col2"), data = df.nj) +
  geom_line(aes(x = time, y = nj, color = "col2"), data = df.nj) +
  geom_point(aes(x = time, y = pa, color = "col1"), data = df.pa) +
  geom_line(aes(x = time, y = pa, color = "col1"), data = df.pa) +
  scale_color_manual(
    name = "Legend",
    values = c("col1" = "red",
               "col2" = "darkgreen"),
    labels = c("Pennsylvania (No Min Wage Increase)", "New Jersey (Min Wage Increase)")
  ) +
  theme_bw() +
  scale_x_continuous(breaks = c(0, 1), labels = c("Mar 1992","Dec 1992")) +
  xlab(NULL) + ylab("Employment Rate") +
  ylim(17, 27)
```

Pennsylvania gives us the **trend** in employment rate if **no minimum wage increase occurred**. We then make an assumption - had New Jersey not undergone a minimum wage increase, it would have followed the same trend in employment rates.

```{r}
#| fig-height: 2
#| echo: false

df.att <- data.frame(
  time = c(1, 1),
  att = c(23.9, 21.1)
)
ggplot() +
  geom_point(aes(x = time, y = nj.pot, color = "col3"), data = df.pot) +
  geom_line(aes(x = time, y = nj.pot, color = "col3"), data = df.pot, linetype = 2) +
  geom_point(aes(x = time, y = nj, color = "col2"), data = df.nj) +
  geom_line(aes(x = time, y = nj, color = "col2"), data = df.nj) +
  geom_point(aes(x = time, y = pa, color = "col1"), data = df.pa) +
  geom_line(aes(x = time, y = pa, color = "col1"), data = df.pa) +
  geom_line(aes(x = time, y = att, color = "col4"), data = df.att, arrow = arrow(length = unit(0.2, "cm"), ends = "both", type = "closed"), linetype = 3) +
  scale_color_manual(
    name = "Legend",
    values = c("col1" = "red",
               "col2" = "darkgreen",
               "col3" = "purple",
               "col4" = "black"),
    labels = c("Pennsylvania (Untreated)", "New Jersey (Treated)", "New Jersey (Hypothetical Untreated)", "Treatment Effect")
  ) +
  theme_bw() +
  scale_x_continuous(breaks = c(0, 1), labels = c("Mar 1992","Dec 1992")) +
  xlab(NULL) + ylab("Employment Rate") +
  ylim(17, 27)
```

Now, this allows us to estimate the counterfactual employment rate of New Jersey had no minimum wage increase occurred. This allows us to find the treatment effect.

This entire idea of using another group's trend to estimate the hypothetical depends on the critical assumption of **parallel trends**: that we can indeed use Pennsylvania's trend to estimate New Jersey's hypothetical trend if it were untreated.

Parallel trends essentially says that if the treated units had hypothetically not been treated, we should expect them to follow the same trend as the untreated units.

::: small
We also need the assumption of SUTVA for difference-in-differences
:::
