---
title: "Gardner (2021) Two-Stage Difference-in-Differences"
---

::: small
See [Kyle Butts'](https://kylebutts.github.io/did2s/articles/Two-Stage-Difference-in-Differences.html) github page for more info on this estimator. [Asjad Naqvi](https://asjadnaqvi.github.io/DiD/docs/code_r/07_did2s_r/) also has a helpful page.
:::

TWFE compares groups with each other - treated vs. not treated, early treated vs. late treated, etc., and these grouped comparisons [can be "forbidden" or incorrect](twfestaggered.qmd).

Gardner (2021) instead proposes to estimate individual counterfactual outcomes, to calculate individual treatment effects, ensuring no forbidden comparisons. Using the idea of two-way fixed effects, we can derive a model for these missing counterfactuals:

$$
\pCred = \unit + \time + \cov + \eps_{it}
$$

::: small
This equation is obtained by taking the [two-way fixed effects](twfe.qmd) equation and plugging in $D_{it} = 0$.
:::

Gardner's 2-stage difference-in-differences uses this equation and framework above, and attempts to estimate $\pCred$.

1.  First, Gardner uses only control units ($D_{it} = 0$) in a regression to estimate $\time$, $\unit$, and $\beta$. The choice of using only control units is to ensure that the estimation of hypothetical controls is accurate and not influenced by the treatment.

$$
Y_{it|D = 0} = \unit + \time + \cov + \eps_{it}
$$

2.  Using estimated $\widehat\unit$, $\widehat\time$, and $\widehat\beta$, we can estimate $\pCred$ counterfactuals for the treated.

$$
\red{\widehat Y_{it}^{(0)}} = \widehat\alpha_i + \widehat\gamma_t + \b X_{it}' \widehat{\b\beta}
$$

3.  Once we have estimated $\red{\widehat Y_{it}^{(0)}}$, we can calculate the treatment effects for the treated units, as we have an estimate for the hypothetical outcomes.

$$
\blue{\widehat\tau_{it}} \approx \pT - \red{\widehat Y_{it}^{(0)}}
$$

4.  Then, we average all the individual treatment effects to get the ATT.

::: small
Steps 3-4 are a simplification - Gardner uses a second regression to calculate the final treatment effects, rather than directly computing the treatment effects with the estimated counterfactual.
:::

<br />

To implement 2-stage difference-in-differences, we need the **did2s** package:

```{r, eval = F}
library(did2s)
```

For the estimation of the ATT, we do the following:

```{r, eval = F}
model <- did2s(data = my_data,
               yname = "outcome",
               first_stage = ~ covariate1 + covariate2 | unit + time,
               second_stage = ~i(treatment),
               treatment = "treatment",
               cluster_var = "unit")
model
```

::: small
If you have no covariates, replace the covariates with 0 (so it looks something like " \~ 0 \| unit + time")
:::

For estimation of dynamic treatment effects (and testing of parallel trends), we do the following:

```{r, eval = F}
model <- did2s(data = my_data,
               yname = "outcome",
               first_stage = ~ covariate1 + covariate2 | unit + time,
               second_stage = ~i(rel_time, ref = c(-1, Inf)),
               treatment = "treatment",
               cluster_var = "unit")
model
iplot(model) #to view dynamic treatment effects plot
```

::: small
If you have no covariates, replace the covariates with 0 (so it looks something like " \~ 0 \| unit + time")

Note: the **inf** is the reference for the never-treated control group, for the variable time_to_treat. If it is NA's, then you do not need to include it, just put ref = -1.
:::
