---
title: "Two-Stage Difference-in-Differences"
subtitle: "Gardner (2021)"
---

```{r}
#| message: false
#| warning: false

# Packages Needed:
library(did2s)
```

The 2-Stage DiD Estimator developed by Gardner (2021) is designed for staggered DiD, accounting for issues in TWFE. The [TWFE estimator](twfe.qmd) implies that untreated units $D_{it} = 0$ equals the following:

$$
Y_{it} (D_{it} = 0) = \unit + \time + \cov + \eps_{it}
$$

::: small
This equation is obtained from plugging in $D_{it} = 0$ into the two-way fixed effects equation.
:::

This means parameters $\unit$, $\time$, and $\beta$ should be estimated with only units $D_{it} = 0$. However, in two-way fixed effects, we estimate $\unit$, $\time$, and $\beta$ with both units $D_{it} = 0$ and $D_{it} = 1$. So, we are actually improperly estimating $Y_{it}(D_{it} = 0)$ in two-way fixed effects.

::: small
This is another fancy way of explaining the [forbidden comparisons problem](twfe.qmd) - in TWFE, $Y_{it}(D_{it} = 0)$ is mis-estimated to include already-treated units, which is the forbidden comparison.
:::

Gardner (2021) proposes to solve this by spliting the TWFE estimator into two parts.

In the [first-stage]{.underline}, we estimate $\unit$, $\time$, and $\beta$ in a regression **with only untreated units** $D_{it} = 0$. This will ensure the proper estimation of $Y_{it}(D_{it} = 0)$. Then, using our first-stage estimates of $\widehat\unit$, $\widehat\time$, and $\widehat\beta$, in the [second stage]{.underline} regression, estimate $\att$.

```{r, echo = F, warning = F, message = F}
data(base_did, package = "fixest")
df <- base_did
colnames(df) <- c("outcome", "covariate", "unit", "time", "post", "ever.treat")
df$treatment <- df$post*df$ever.treat
```

```{r}
#| warning: false
#| message: false
#| comment: "#>"
#| class-output: r

model <- did2s(
  data = df,
  yname = "outcome",
  first_stage = ~ covariate | unit + time,
      # if no covariate, should be ~ 0 | unit + time
  second_stage = ~ i(treatment), #replace D with your treatment
  treatment = "treatment",
  cluster_var = "unit")
summary(model)
```

::: small
Our ATT is the estimate for treatment::1, which in this case is 4.845.
:::

For estimation of dynamic treatment effects, we do the following:

```{r}
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
first.year <- df %>%
  filter(treatment == 1) %>%
  group_by(unit) %>%
  summarise(first.treat = min(time))

df <- left_join(df, first.year)
df$rel.time <- df$time - df$first.treat

df$rel.time <- replace_na(df$rel.time, 10000)
```

```{r}
#| warning: false
#| message: false
#| fig-height: 3.5

model <- did2s(
  data = df,
  yname = "outcome",
  first_stage = ~ covariate | unit + time,
      # if no covariate, should be ~ 0 | unit + time
  second_stage = ~ i(rel.time, ref = c(-1, 10000)),
      # the first number in ref= should always be -1
      # the second number in ref= should be the value of rel.time for the never-treated
  treatment = "treatment",
  cluster_var = "unit")

iplot(
  model,
  drop = "[[:digit:]]{2}",
      # this drops the time periods too far before and after treatment.
  ref.line = 0)
```

<br />

::: small
See [Kyle Butts'](https://kylebutts.github.io/did2s/articles/Two-Stage-Difference-in-Differences.html) github page for more info on this estimator. [Asjad Naqvi](https://asjadnaqvi.github.io/DiD/docs/code_r/07_did2s_r/) also has a helpful page.

Also see the [documentation](https://kylebutts.github.io/did2s/reference/did2s.html) of the did2s function.
:::
