---
title: "Dynamic Treatment Effects"
---

::: small
This page requires knowledge on [two-way fixed effects](twfe.qmd).
:::

The classic difference-in-differences design only requires 2 time periods - a pre-treatment periods, and a post-treatment period. However, we often have more data - many years before the treatment began, and many years after the treatment ended.

**Dynamic Treatment Effects** calculate treatment effects for each year before treatment, and each year after treatment, instead of just one ATT.

::: small
This is also called an **event study** or **leads-and-lags**. Just like TWFE, this estimator has [issues when dealing with staggered treatment](twfestaggered.qmd).
:::

1.  We first create a new categorical variable, called relative-time $R_{it}$. It essentially measures the number of periods a certain time period $t$ is before/after the initial treatment adoption, with $R_{it} = 0$ being the initial treatment year.

$$
R_{it} = t - \text{initial treatment period}
$$

::: small
For example, if treatment starts in 2003, an observation in 2002 would get a relative-time of -1, and an observation in 2004 would get a relative-time of 1.
:::

2.  Then, we run the following model:

$$
Y_{it} = \underbrace{\unit + \time}_{\text{fixed effects}} + R_{it}\blue{\tau_{r}}+ \cov + \eps_{it}
$$

::: small
$R_{it}$ is a categorical variable, with $R_{it} = -1$ set as the reference category (since this is the last pre-treatment period). If you have repeated cross-sections, use group-fixed effects instead of unit.
:::

3.  This will estimate a coefficient $\blue{\tau_{r}}$ for each relative time period before/after treatment implementation. These are the treatment effects for each relative-time period $r$.

::: small
More technically, $\tau_r$ is the estimated difference in $Y_{it}$ compared to reference periods $R_{it} = -1$ (which is assumed to have a 0 treatment effect).
:::

In the **post-treatment** periods (after treatment has started), these $\blue{\tau_r}$ represent the treatment effect by year. We can graph the post-treatment $\blue{\tau_r}$ to see the trend in how the effect of treatment changes over time.

In the **pre-treatment** periods, we shouldn't expect any singificant causal effect $\blue{\tau_r}$, since treatment hasn't started. If we do see any statistically significant effect pre-treatment, [this is evidence for a violation of the parallel trends assumption]{.underline}.

::: small
Note: this is not a definitive test for parallel trends. Even if there is 0 significance in pre-treatment periods, parallel trends could still be violated. There is no way to be 100% sure.
:::

<br />

To implement dynamic treatment effects, we need the **fixest** package:

```{r, eval = F}
library(fixest)
```

Before we start, we will need to create a relative-time variable. This can be done by taking the time variable, and subtracting the initial treatment time period for each unit.

Then, we run the following model:

```{r, eval = F}
model <- feols(outcome ~ i(relative.time, -1) | unit + time,
               data = my.data,
               se = "cluster")
iplot(model)
```

::: small
If you are using repeated cross-section, replace unit with group fixed effects.
:::

The output from **iplot()** will be a graph with each $\blue{\tau_r}$ and its confidence interval. If any of the confidence intervals do not touch 0, then it is statistically significant.
