---
title: "Extended Two-Way Fixed Effects"
subtitle: "Wooldridge (2021, 2023)"
---

```{r}
#| message: false
#| warning: false

# Packages Needed
library(etwfe)
```

The Extended Two-way Fixed Effects Estimator developed by Wooldridge (2021, 2023) is designed for staggered DiD, accounting for issues in TWFE. It is also capable of dealing with non-linear models such as negative binomial, poisson, and logistic.

The ETWFE estimator estimates all possible heterogeneous effects separately (by including a lot of interactions with initial treatment year, year, and covariates), using only valid comparisons, solving the forbidden comparison problem:

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
data(base_stagg, package = "fixest")
df <- base_stagg
colnames(df) <- c("unit", "time", "first.treated", "rel.time", "treat", "te", "covariate", "outcome")
```

```{r}
#| warning: false
#| message: false

model <- etwfe(
  fml = outcome ~ covariate,
  tvar = time,
  gvar = first.treated,
  data = df,
  vcov = ~unit) 
```

Many coefficients $\blue{\tau_{it}}$ and $\blue{\tau_{itx}}$ are estimated that capture different heterogeneity in treatment effects. These coefficients are manually aggregated together into one ATT using proper weighting based on how frequent each heterogeneous treatment effect occurs in our sample. This solves the weighting problem of TWFE.

```{r}
#| comment: "#>"
#| class-output: r
emfx(model)
```

::: small
We can see our estimated ATT is -1.21.
:::

We can also aggregate these coefficients into dynamic treatment effects. However, ETWFE only supports post-treatment dynamic treatment effects, not pre-treatment effects.

```{r}
#| fig-height: 3.5

dynamic <- emfx(model, type = "event")
plot(dynamic)
```

::: small
Since ETWFE does not support pre-treatment effects, we cannot use it to test for parallel trends.
:::

Wooldridge's 2023 paper also extends this framework to work with logistic regression, poisson regression, and negative binomial regression, which is useful when we deal with non-continuous outcome variables. This can be accessed by adding an argument to the **etwfe()** function:

```{r, eval = F}
family = c("logit", "poisson", "negbin")
```

::: small
Only choose one of them - you cannot choose multiple families together.
:::

<br />

::: small
For technical details, see [Wooldridge (2021)](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3906345) and [Wooldridge (2023)](https://academic.oup.com/ectj/article/26/3/C31/7250479).

For details on the etwfe function, see the [documentation](https://grantmcdermott.com/etwfe/reference/index.html).
:::
