---
title: "Imputation Estimator"
subtitle: "Liu, Xu, and Wang (2024)"
---

```{r}
#| message: false
#| warning: false

# Packages Needed:
library(fect)
```

The Imputation Estimator developed by Liu, Xu, and Wang (2024) is designed for staggered DiD and Non-Absorbing DiD (treatment reversal), accounting for issues in TWFE.

The estimator proposes to estimate individual counterfactual untreated outcomes for treated units, and directly calculating the individual treatment effects:

$$
\tau_{it} = \pT - \pCred
$$

Thus, the goal of this **interactive fixed effects counterfactual estimator (IFEct)** is to estimate $\pCred$. They model these counterfactuals as:

$$
\pCred = \underbrace{\unit + \time + \cov}_{\text{classic TWFE}} + \b\lambda_i'\b\xi_t + \eps_{it}
$$

::: small
$\b\xi_t$ is a vector of latent variables/factors $\xi_{1t}, \xi_{2t}, \dots$. These latent variables change in magnitude depending on time $t$ (they are time-varying). $\b\lambda_i$ is a vector of factor loadings $\lambda_{1i}, \lambda_{2i}, \dots$, that describes the "influence" every latent factor has on unit $i$. Different units $i$ can have different relationships with each factor.
:::

```{r, echo = F, warning = F, message = F}
library(fect)
set.seed(2003)
df <- did2s::gen_data(
        panel = c(1990, 2020),
        g1 = 2000, g2 = 2010, g3 = 0,
        te1 = 2, te2 = 1, te3 = 0,
        te_m1 = 0.05, te_m2 = 0.15, te_m3 = 0,
        n=100)
colnames(df) <- c("unit", "state", "group",
                  "unit.fe", "first.treat", "time", "year.fe",
                  "treatment", "rel.time", "rel.time.binned", "error",
                  "te", "te_dynamic", "outcome")
df$treatment <- as.numeric(df$treatment)
```

```{r}
#| warning: false
#| message: false
#| comment: "#>"
#| class-output: r


set.seed(17) #for consistent replication
model <- fect(
  formula = outcome ~ treatment,
      #add covariates after treatment
  data = df,
  index = c("unit","time"),
      #replace unit and time with your unit and time variables
  method = "ife",
      #do not change
  se = TRUE,
  nboots = 50) #for standard errors. more is more accurate, but takes longer
print(model)
```

We have two outputs - both are plausible ATT's. It is generally recommended to use the first row - treated observations equally weighted, as it is typically the most accurate ATT estimate.

We can also plot our results for dynamic treatment effects and parallel trends tests:

```{r}
#| fig-height: 3.2
plot(model)
```

::: small
See [Liu et al (2024)](https://onlinelibrary.wiley.com/doi/10.1111/ajps.12723) for more technical details.
:::
