---
title: "PanelMatch Estimator"
subtitle: "Imai, Kim, and Wang (2023)"
---

```{r}
#| message: false
#| warning: false

# Packages Needed:
library(PanelMatch)
```

The PanelMatch estimator developed by Imai, Kim, and Wang (2023) is designed for staggered DiD and Non-absorbing DiD (treatment reversal), accounting for issues in TWFE. Before we start, we need to clean our data.

```{r}
#| echo: false
#| message: false
#| warning: false

data(base_did, package = "fixest")
df <- base_did
colnames(df) <- c("outcome", "covariate", "unit", "time", "post", "ever.treat")
df$treat <- df$post*df$ever.treat
```

```{r}
#| message: false
#| warning: false

# we need unit and time var to be integers
df$unit <- as.integer(df$unit)
df$time <- as.integer(df$time)

# PanelMatch does not like tidy dataframes, so do this:
df <- as.data.frame(df)

df.cleaned <- PanelData(
  panel.data = df,        # your data
  unit.id    = "unit",    # your unit var (integer only)
  time.id    = "time",    # your time period var (integer only)
  treatment  = "treat",   # your treatment var
  outcome    = "outcome"  # your outcome var
)
```

The estimator is based on the potential outcomes framework. For a treated unit, their individual causal effect is.

$$
\tau_{it} = \pT - \pCred
$$

Our issue is that we do not observe $\pCred$ for treated units. PanelMatch finds an untreated unit from the same time period, with similar "history" and characteristics, and uses that unit as the missing potential outcome:

$$
\tau_{it} \approx \pT - \red{Y_{\text{matched individual}, t}^{(0)}}
$$

How does PanelMatch determine which untreated unit is selected as a match? It depends on 2 things:

1.  The past "history" of the matched individual before time period $t$. This includes their past treatment history, and their past outcome values.

::: small
The authors consider the history in the "lag" period - which is a set number of periods before the given $t$ (we can choose how many years).
:::

2.  (optional) if covariates are specified, it will use also consider distance matching on covariates to find a match.

```{r}
#| message: false
#| warning: false

match <- PanelMatch(
  lag               = 3,              # how many periods of history to consider
  panel.data        = df.cleaned,     # PanelData object
  lead              = c(0:3),         # how many post-treat dynamic effects
  covs.formula      = ~ covariate,    # covaraites to include
  refinement.method = "mahalanobis",  # if no covariates, put "none"
  qoi               = "att",
  match.missing     = TRUE
)
```

Once we have found the matches, we can now estimate the ATT:

```{r}
#| message: false
#| warning: false
#| comment: "#>"
#| class-output: r

model <- PanelEstimate(
  sets       = match,       # PanelMatch object
  panel.data = df.cleaned,  # PanelData object
  pooled     = TRUE
)

summary(model)
```

The output will be our ATT estimate. We can also estimate dynamic treatment effects for post-treatment periods:

```{r}
#| message: false
#| warning: false
#| comment: "#>"
#| class-output: r

model <- PanelEstimate(
  sets       = match,       # PanelMatch object
  panel.data = df.cleaned,  # PanelData object
  pooled     = FALSE
)

summary(model)
```
