---
title: "Sun and Abraham (2021) Estimator"
---

```{r}
#| message: false
#| warning: false

# Packages Needed:
library(fixest)
```

There are two [issues with the two-way fixed effects estimator](staggered.qmd): forbidden comparisons and issues with weighting. Sun and Abraham (2021) proposes a new estimator to solve this issues.

The estimator divides all units into groups based on their initial treatment implementation. These groups are encoded into a new categorical variable $G_i$.

Then, the estimator calculates dynamic ATT's separately for group, using only valid comparisons. This is done by interacting a relative-time variable $R_{it}$ with $G_i$. These interactions will produce many coefficients $\blue{\tau_{it}}$ (in fact, one for each time period for each initial treatment year group). These coefficients capture all the heterogeneity in effects.

These coefficients are manually aggregated together into dynamic ATT's by relative treatment time using proper weighting. This solves the weighting problem with TWFE.

```{r}
#| echo: false
#| warning: false
#| message: false

data(base_stagg, package = "fixest")
df <- base_stagg
colnames(df) <- c("unit", "time", "first.treated", "rel.time", "treat", "te", "covariate", "outcome")
```

```{r}
#| warning: false
#| message: false
#| fig-height: 3.5

model <- feols(
  outcome ~ sunab(first.treated, time) + covariate | unit + time,
  data = df,
  vcov = ~unit)

iplot(
  model,
  drop = "[[:digit:]]{2}",
      # this drops the time periods too far before and after treatment.
  ref.line = 0)
```

We can also aggregate the dynamic ATT's into one singular ATT:

```{r}
#| warning: false
#| message: false
#| comment: "#>"
#| class-output: r

aggregate(model, "att")
```

<br />

::: small
For technical details, see [Sun and Abraham (2021)](https://www.sciencedirect.com/science/article/abs/pii/S030440762030378X) or [this github page](https://asjadnaqvi.github.io/DiD/docs/code_r/07_sunab_r/).

For more options with the sunab and aggregate functions, see the [documentation for fixest](https://lrberge.github.io/fixest/reference/index.html).
:::
