---
title: "Two-Way Fixed Effects Estimator"
---

::: small
Note: if you have staggered treatment implementation, [TWFE can produce incorrect results](twfestaggered.qmd).
:::

The two-way fixed effects estimator is the most commonly used estimator in difference-in-differences, used when we have panel data with multiple observations observed in multiple time periods. The estimator is a linear regression model specified as following:

$$
\Y = \underbrace{\unit + \time}_{\text{fixed effects}} + \blue\tau \ \D + \beta\ \cov + \eps
$$

::: small
Where covariates are control variables needed to meet the parallel trends assumption (optional).
:::

The fixed effects $\alpha_i$ and $\gamma_t$ are "special" intercepts. Unlike our normal intercepts in regression, these intercepts $\alpha_i$ and $\gamma_t$ take different values depending on the unit and time.

-   Intercept $\alpha_i$ has a different value for each unit $i$.
-   Intercept $\gamma_t$ has a different value for each time period $t$.

Essentially, what fixed effects does is that it calculates different intercepts for each unit $i$, and each time period $t$. Thus, these intercepts reflect differences on-average between different units $i$, and different time periods $t$.

Since the differences between units $i$ and time periods $t$ are included in these fixed effects intercept terms, this allows us to **account** for differences between units and differences between time periods.

::: small
Or in other words, we are essentially "controlling" for differences between units, and differences between years.
:::

By accounting for differences between units and between differences in time periods, we allow our observations to be more on a level playing field, allowing for us to calculate the causal effect. The estimate for $\blue\tau$ is our treatment effect (ATT). This estimate is **unbiased** if the parallel trends assumption is met.

<br />

To implement two-way fixed effects, we need the **fixest** package:

```{r, eval = F}
library(fixest)
```

We implement two-way fixed effects as follows:

```{r, eval = F}
model <- feols(outcome ~ treatment + covariate1 + covariate2 | unit + time,
               data = my_data,
               se = "cluster")
```
