---
title: "Two-Way Fixed Effects Estimator"
---

The two-way fixed effects estimator is the standard estimator for DiD.

::: small
Note: if you have staggered treatment implementation, [TWFE can produce incorrect results](twfestaggered.qmd).
:::

The estimator with **panel data** is a linear regression model specified as following:

$$
Y_{it} = \underbrace{\unit + \time}_{\text{fixed effects}} + D_{it}\att + \cov + \eps_{it}
$$

::: small
Panel data is when we observe the same units $i$ over multiple periods of time.
:::

For **repeated cross-section** data, the model is:

$$
Y_{it} = \underbrace{\alpha_g + \time}_{\text{fixed effects}} + D_{it}\att + \cov + \eps_{it}
$$

::: small
Repeated cross section is when we have different samples from different time periods. Note how the first fixed effect $\alpha_g$ now has subscript $g$ - we now include group fixed effects - group being the level on which treatment is assigned (ex. if treatment is assigned at a state level, we use state fixed effects).
:::

The fixed effects $\alpha_i$ and $\gamma_t$ are "special" intercepts. Unlike our normal intercepts in regression, these intercepts take different values depending on the unit and time.

-   Intercept $\alpha_i$ has a different value for each unit $i$.
-   Intercept $\gamma_t$ has a different value for each time period $t$.

::: small
For repeated cross section, $\gamma_t$ is the same. $\alpha_g$ will now have different values for each group $g$.
:::

What fixed effects does is that it calculates different intercepts for each unit $i$, and each time period $t$. Thus, these intercepts reflect differences on-average between different units $i$, and different time periods $t$.

Since the differences between units $i$ and time periods $t$ are included in these fixed effects intercept terms, this allows us to **account** for differences between units and differences between time periods.

::: small
Or in other words, we are essentially "controlling" for differences between units, and differences between years.
:::

By accounting for differences between units and between differences in time periods, we allow our observations to be more on a level playing field, allowing for us to calculate the causal effect. The estimate for $\att$ is our treatment effect (ATT). This estimate is **unbiased** if the parallel trends assumption is met.

<br />

To implement two-way fixed effects, we need the **fixest** package:

```{r, eval = F}
library(fixest)
```

We implement two-way fixed effects as follows:

```{r, eval = F}
model <- feols(outcome ~ treatment + covariate1 + covariate2 | unit + time,
               data = my_data,
               se = "cluster")
summary(model)
```

::: small
For repeated cross section, replace **unit** with **group**, group being the level at which treatment is assigned.
:::
