---
title: "Fully Interacted Estimator"
---

::: small
See the pros and cons of this estimator in the [choosing an estimator](soochoose.qmd) page.
:::

From [selection on observables](soo.qmd), we know that our causal effect is a weighted average:

$$
\ate = \sum\tau_\text{X} \cdot Pr( X)
$$

Notice how the weights are the probability of the confounder values of $\b x$. With some complex math (Angrist 1998), we can actually show that OLS actually estimates:

$$
\hat\beta_\text{OLS} = \sum \tau_X\cdot \underbrace{\frac{Var(D_i | X)Pr(X)}{\sum Var(D_i | X^c)Pr( X^c)}}_{\text{weight}}
$$

::: small
Where $X^c$ is the complement (not $X$).
:::

These weights are not equivalent to the selection on observables $\ate$. Thus, if not all $\tau_\text{X}$ are exactly the same (which implies heterogeneity), then our linear regression estimator will incorrectly estimate the ATE.

Heterogeneity is present in almost all situations we are interested in. [Lin (2013)](https://projecteuclid.org/journals/annals-of-applied-statistics/volume-7/issue-1/Agnostic-notes-on-regression-adjustments-to-experimental-data--Reexamining/10.1214/12-AOAS583.full) proposes the **fully interacted estimator**, which allows for consistent estimation of the ATE even with heterogeneity:

$$
Y_i = \alpha + D_i\ \ate + \underbrace{(\b X_i - \b{\mean X})'\b\beta + D_i(\b X_i - \b{\mean X})'\b\gamma}_{\text{interactions with de-meaned covariates}}+ \eps_i
$$

The new OLS estimate of $\ate$ in this estimator will technically still be a biased estimator of the ATE, but the bias is negligible.

<br />

Before you implement the estimator, make sure you have reasons to believe you meet the neccessary assumptions for [selection on observables](soo.qmd).

We will need the **estimatr** package:

```{r, eval = FALSE}
library(estimatr)
```

Then, we can use the **lm_lin()** function to estimate:

```{r, eval = FALSE}
ate <- estimatr::lm_lin(outcome ~ treatment,
                        covariates = ~ covariate1 + covariate2,
                        data = my_data)
summary(ate)
```

The output will be the ATE - the average treatment effect for all units in the study.

<br />
