---
title: "Propensity Score Matching"
---

```{r}
#| message: false
#| warning: false

# Packages Needed:
library(Matching)
```

Mia is in our study and receives the treatment. Mia's causal effect is:

$$
\tau_{\text{Mia}} = \textcolor{green}{Y^{(1)}_\text{Mia}} - \textcolor{red}{Y^{(0)}_\text{Mia}}
$$

We cannot observe Mia's counterfactual (in red), since Mia receives the treatment.. However, what we can do is to find an untreated individual similar to Mia to approximate Mia's counterfactual:

$$
\tau_{\text{Mia}} \approx \textcolor{green}{Y^{(1)}_\text{Mia}} - \textcolor{red}{Y^{(0)}_\text{Matched Individual}}
$$

Propensity Score Matching matches an individual that is treated (like Mia) with one that is not treated based on how similar their likelihoods of treatment are.

What is a likelihood of treatment? Well we know confounders cause people to get the treatment or not treatment. Thus, using an individual's confounder values, we can estimate their likelihood of getting treatment, called a propensity score.

$$
\text{propensity score } \pi =Pr(\green{D_i = 1}) 
$$

Propensity scores are typically estimated with a logistic regression.

::: small
This also means that propensity score matching shares the same weaknesses of logistic regression - including assuming linear relationships between confounders and propensities, and only being unbiased in large sample sizes.
:::

```{r}
#| echo: false
#| message: false
#| warning: false


library(tidyverse)
N = 1000
ATE = 2
  
df = tibble(
  cov1 = rbinom(N, 1, 0.5),
  cov2 = rnorm(N),
  P = rnorm(N),
  U = rnorm(N), 
  W = rnorm(N),
) %>%
  mutate(
    Y0 = 1 + 0.5*cov1 + 0.5*cov2 + 0.5*P + 0.5*U,
    Y1 = Y0 + rnorm(N, mean = ATE*1.5, sd = 1)*cov1 + rnorm(N, mean = ATE/2)*(1-cov1)
  )

df = df %>%
  mutate(
    U2 = rnorm(N),
    pscore = 1/(1 + exp(-(-1 - 2*cov1 - 1.5*cov2 + 1.5*W + 0.5*U2))),
    treatment = rbinom(N, 1, pscore),
    outcome = Y1*treatment + Y0*(1-treatment)
  )
```

```{r}
#| warning: false
#| message: false
#| comment: "#>"
#| class-output: r

# Logistic Model for Propensity Scores
propensity <- glm(
  treatment ~ cov1 + cov2,
  data = df,
  family = "binomial")

# Predice Propensity Scores for All Observations
df$pscore <- predict(propensity, type = "response")
```

Now with the propensity scores, we can conduct [distance matching](distance.qmd), matching units by how close their propensity scores are.

```{r}
#| warning: false
#| message: false
#| comment: "#>"
#| class-output: r

att <- Match(
  Y = df$outcome,
  Tr = df$treatment,
  X = df[,"pscore"],
  M = 1,
      # number of matches - default is 1 to 1, but we can match more to 1
  BiasAdjust = TRUE,
      # modern adjustment - usually good to keep on
  Weight = 2)
summary(att)
```

Our output estimate will be the ATT - the average treatment effect for those units who received the treatment.

+----------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------+
| Pros                                                                                                                                         | Cons                                                                                                                                                  |
+==============================================================================================================================================+=======================================================================================================================================================+
| 1\) Non-parametric, so it does not need to assume a linear relationship between confounders and outcome like the fully interacted estimator. | 1\) Requires a large sample size to be unbiased, because logistic regression for the propensity scores does not work well with small amounts of data. |
|                                                                                                                                              |                                                                                                                                                       |
| 2\) Can accurately handle more confounders than distance matching.                                                                           | 2\) Throws out unmatched data, so it wastes data.                                                                                                     |
|                                                                                                                                              |                                                                                                                                                       |
|                                                                                                                                              | 3\) Does not estimate the ATE, only estimates the ATT.                                                                                                |
+----------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------+

: {tbl-colwidths="\[50,50\]"}

::: small
Kevin's Comments: there is little reason to use propensity score matching over genetic matching, unless your machine cannot estimate genetic matching.
:::
