---
title: "Linear Regression Estimator"
---

::: small
See the pros and cons of this estimator in the [choosing an estimator](soochoose.qmd) page.
:::

Selection on observables is about controlling for confounders. Linear regression is a very natural way to control for confounders.

$$
\Y = \beta_0 + \beta_1\ \X_{1} + \beta_2\ \X_{2} + \eps
$$

::: small
Where $\X_1$ and $\X_2$ are two different input/explanatory/independent variables
:::

We know that in the above equation, $\beta_1$ is the relationship between $\X_1$ and outcome, while controllling (holding constant) $\X_2$.

Using this idea, we can implement causal inference with regression, with one of the explanatory variables being our treatment variable, and the rest of the explanatory variables being control variables.

$$
\Y = \alpha + \blue\tau\ \D + \beta\ \cov + \eps
$$

::: small
Where $\alpha$ is the intercept and $\cov_i$ is a vector of covariates/confounders for individual $i$.
:::

Our ordinary least squares (OLS) estimate $\blue{\widehat\tau}$ is an unbiased estimator of the true $\tau_{ATE}$ given three conditions are met:

1.  We meet the selection on observables assumption of **conditional ignorability**. Conditional ignorability implies exogeneity, which means the estimate is unbiased.

2.  The relationship between our continuous confounders and outcome variable is **linear**. This is because if the true relationship between these two is not linear, then our linear model is wrong, so it is not properly controlling for confounders.

3.  There is **no heterogeneity** in treatment effects. Angrist (1998), Lin (2013), and Słoczyński (2022) have proven that when there is heterogeneity, OLS is estimating another quantity that is not the ATE.

::: small
**Heterogeneity** means that different individuals have different individual treament effects $\tau_i$. OLS only estimates the ATE if there is homogeneity - all $\tau_i$ are equal.
:::

<br />

Before you implement the estimator, make sure you have reasons to believe you meet the neccessary assumptions for [selection on observables](soo.qmd).

We will need the **fixest** package:

```{r, eval = F}
library(fixest)
```

To run a regression estimator, we do:

```{r, eval = F}
feols(outcome ~ treatment + covariate1 + covariate2,
      data = data,
      se = "hetero")
```

::: small
We typically assume heteroscedasticity, so we use heteroscedasticity-robust standard errors. If you can prove homoscedasticity, then you can use normal standard errors.
:::

The coefficient for the treatment variable will be the ATE - the average treatment effect for all units in the study.

::: small
Assuming you have met all the assumptions of selection on observables, and the special assumptions for the linear regression estimator shown above.
:::
