---
title: "Linear Regression Estimator"
---

::: small
See the pros and cons of this estimator in the [choosing an estimator](soochoose.qmd) page.
:::

Selection on observables is about controlling for confounders. Linear regression is a very natural way to control for confounders.

Using this idea, we can implement causal inference with regression, with one of the explanatory variables being our treatment variable, and the rest of the explanatory variables being control variables.

$$
Y_i = \alpha + D_i\ \ate + \covs + \eps_i
$$

Our ordinary least squares (OLS) estimate $\hat\ate$ is an unbiased estimator of the true ATE given three conditions are met:

1.  We meet the selection on observables assumption of **conditional ignorability**. Conditional ignorability implies exogeneity, which means the estimate is unbiased.

2.  The relationship between our continuous confounders and outcome variable is **linear**. This is because if the true relationship between these two is not linear, then our linear model is wrong, so it is not properly controlling for confounders.

3.  There is **no heterogeneity** in treatment effects. Angrist (1998), Lin (2013), and Słoczyński (2022) have proven that when there is heterogeneity, OLS is estimating another quantity that is not the ATE. This is covered in more detail on [fully interacted estimator](interact.qmd) page

::: small
**Heterogeneity** means that different individuals have different individual treament effects $\tau_i$. OLS only estimates the ATE if there is homogeneity - all $\tau_i$ are equal.
:::

<br />

Before you implement the estimator, make sure you have reasons to believe you meet the neccessary assumptions for [selection on observables](soo.qmd).

We will need the **fixest** package:

```{r, eval = F}
library(fixest)
```

To run a regression estimator, we do:

```{r, eval = F}
feols(outcome ~ treatment + covariate1 + covariate2,
      data = data,
      se = "hetero")
```

::: small
We typically assume heteroscedasticity, so we use heteroscedasticity-robust standard errors. If you can prove homoscedasticity, then you can use normal standard errors.
:::

The coefficient for the treatment variable will be the ATE - the average treatment effect for all units in the study.

::: small
Assuming you have met all the assumptions of selection on observables, and the special assumptions for the linear regression estimator shown above.
:::
