---
title: "Distance Matching"
---

```{r}
#| message: false
#| warning: false

# Packages Needed:
library(Matching)
```

Mia is in our study and receives the treatment. Mia's causal effect is:

$$
\tau_{\text{Mia}} = \textcolor{green}{Y^{(1)}_\text{Mia}} - \textcolor{red}{Y^{(0)}_\text{Mia}}
$$

We cannot observe Mia's counterfactual (in red), since Mia receives the treatment.. However, what we can do is to find an untreated individual similar to Mia to approximate Mia's counterfactual:

$$
\tau_{\text{Mia}} \approx \textcolor{green}{Y^{(1)}_\text{Mia}} - \textcolor{red}{Y^{(0)}_\text{Matched Individual}}
$$

Distance matching matches an individual that is treated (like Mia) with one that is not treated based on how **close** their confounding values are. We define closeness by Mahalanobis distance:

$$
\text{distance}_{i, j} = \sqrt{(\b x_i - \b x_j)' \ \b\Sigma_x^{-1} (\b x_i - \b x_j)}
$$

::: small
Where $i$ and $j$ are two units we want to measure the distance between, $\b x$ is a vector of confounder values, and $\b\Sigma_x$ is the covariance matrix of confounders. We must include **all confounders**.
:::

```{r}
#| echo: false
#| message: false
#| warning: false


library(tidyverse)
N = 1000
ATE = 2
  
df = tibble(
  cov1 = rbinom(N, 1, 0.5),
  cov2 = rnorm(N),
  P = rnorm(N),
  U = rnorm(N), 
  W = rnorm(N),
) %>%
  mutate(
    Y0 = 1 + 0.5*cov1 + 0.5*cov2 + 0.5*P + 0.5*U,
    Y1 = Y0 + rnorm(N, mean = ATE*1.5, sd = 1)*cov1 + rnorm(N, mean = ATE/2)*(1-cov1)
  )

df = df %>%
  mutate(
    U2 = rnorm(N),
    pscore = 1/(1 + exp(-(-1 - 2*cov1 - 1.5*cov2 + 1.5*W + 0.5*U2))),
    treatment = rbinom(N, 1, pscore),
    outcome = Y1*treatment + Y0*(1-treatment)
  )
```

```{r}
#| warning: false
#| message: false
#| comment: "#>"
#| class-output: r

model <- Match(
  Y = df$outcome,
  Tr = df$treatment,
  X = df[,c("cov1","cov2")],
  M = 1,
      # number of matches - default is 1 to 1, but we can match more to 1
  BiasAdjust = TRUE,
      # modern adjustment - usually good to keep on
  Weight = 2)
summary(model)
```

If [all confounders are included]{.underline}, our output estimate will be the ATT - the average treatment effect for those units who received the treatment.

::: small
Omission of any confounder will cause inaccurate matches, thus causing inaccurate results.
:::

Distance matching suffers from the curse of dimensionality. This means that as the number of confounders increase, the bias in distance matching estimates becomes larger. Thus, we should avoid using distance matching with more than 3-5 confounders.

::: small
Kevin's comments: there is little reason to use distance matching over [genetic matching](genetic.qmd), unless your machine lacks the computational power for genetic matching.
:::

Distance matching (and all matching) also throws out a lot of data that is unmatched, which can be wasteful.
