---
title: "Fully Interacted Estimator"
---

```{r}
#| message: false
#| warning: false

# Packages Needed:
library(estimatr)
```

Let us say we have data on all confounders. A natural way to account for all confounders is to control for all confounders in a linear regression. With all confounders included, our model should be exogenous, and the $\beta_\text{OLS}$ estimate should be unbiased.

$$
Y_i = \alpha + D_i\beta_\text{OLS} + \covs
$$

However, there is an issue with OLS estimates under treatment heterogeneity. We know that our causal effect is a weighted average of our conditional treatment effects:

$$
\ate = \sum\tau_\text{CATE}(x) \cdot Pr( x)
$$

Notice how the weights are the probability of the confounder values of $\b x$. With some complex math (Angrist 1998), we can show OLS estimates:

$$
\hat\beta_\text{OLS} = \sum \tau_\text{CATE}(x)\cdot \underbrace{\frac{Var(D_i | x)Pr(x)}{\sum Var(D_i | x^c)Pr( x^c)}}_{\text{weight}}
$$

These weights are not equivalent to the selection on observables $\ate$. Thus, if not all $\tau_\text{X}$ are exactly the same (homogeneity), then our OLS estimator (OLS) will incorrectly estimate the ATE. [Lin (2013)](https://projecteuclid.org/journals/annals-of-applied-statistics/volume-7/issue-1/Agnostic-notes-on-regression-adjustments-to-experimental-data--Reexamining/10.1214/12-AOAS583.full) proposes the **fully interacted estimator**, which allows for consistent estimation of the ATE even with heterogeneity:

$$
Y_i = \alpha + D_i\ \ate + \underbrace{(\b X_i - \b{\mean X})'\b\beta + D_i(\b X_i - \b{\mean X})'\b\gamma}_{\text{interactions with de-meaned covariates}}+ \eps_i
$$

```{r}
#| echo: false
#| message: false
#| warning: false


library(tidyverse)
N = 1000
ATE = 2
  
df = tibble(
  cov1 = rbinom(N, 1, 0.5),
  cov2 = rnorm(N),
  P = rnorm(N),
  U = rnorm(N), 
  W = rnorm(N),
) %>%
  mutate(
    Y0 = 1 + 0.5*cov1 + 0.5*cov2 + 0.5*P + 0.5*U,
    Y1 = Y0 + rnorm(N, mean = ATE*1.5, sd = 1)*cov1 + rnorm(N, mean = ATE/2)*(1-cov1)
  )

df = df %>%
  mutate(
    U2 = rnorm(N),
    pscore = 1/(1 + exp(-(-1 - 2*cov1 - 1.5*cov2 + 1.5*W + 0.5*U2))),
    treatment = rbinom(N, 1, pscore),
    outcome = Y1*treatment + Y0*(1-treatment)
  )
```

```{r}
#| warning: false
#| message: false
#| comment: "#>"
#| class-output: r

model <- lm_lin(
  outcome ~ treatment,
  covariates = ~ cov1 + cov2,
  data = df)
fixest::coeftable(model)[,1:4] # Can also use summary(model)
```

::: small
The estimate for treatment is our ATE estimate.
:::

If [all confounders are included]{.underline}, the new OLS estimate of $\ate$ in this estimator will technically still be a biased estimator of the ATE, but the bias is negligible.

This estimator (as a linear estimator naturally does) assumes that the relationship between confounders and outcome is linear. If this assumption is hard to justify, then our estimate of $\ate$ will be biased.
